---
title: "Bioassessment Dashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    source_code: embed
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(flexdashboard)
library(crosstalk)
library(plotly)
library(DT)
library(htmltools)
library(lubridate)
library(summarywidget)
library(sf)
library(leaflet)
library(inlmisc)

```

Sidebar {.sidebar}
=====================================

### Dataset Filters

Choose a Collector ID, Basin, and StationID to plot data on subsequent tabs. Note you can see the table to the right update as you progressively filter the dataset. 
```{r sidebar, echo=FALSE}
IR_SCI <- readRDS('IR_SCI2.RDS')
IR_SCI_HAB2 <- readRDS('IR_SCI3.RDS')
# SCI Table
IR_SCI_Table <- select(IR_SCI, BenSampID:SCI, method, FamTotTaxa:VCPMI65) 
IR_SCI_Table$CollDate <- as.Date(IR_SCI_Table$CollDate)
# Habitat data for plotly
IR_HAB <- select(IR_SCI, BenSampID, StationID, CollDate, CollectorID, Season, Year, HabSampID:TotalHabitat) %>%
  mutate(minX = min(CollDate)- months(1), maxX = max(CollDate) + months(1), 
         hab0 = 0, hab1 = 100, hab2 = 130, hab3 = 150, hab4= 200)
# Habitat data for heatmap
IR_HAB_heatmap <- select(IR_HAB, BenSampID, HabSampID, CollDate, EnterDate, Comments, ALTER:TotalHabitat) %>%
  filter(!is.na(TotalHabitat))
IR_HAB_heatmap$CollDate <- as.Date(IR_HAB_heatmap$CollDate)
IR_HAB_heatmap$EnterDate <- as.Date(IR_HAB_heatmap$EnterDate)
# SCI Raw
IR_SCI_Raw <- IR_SCI %>%  
  group_by(Year, RepNum) %>%
  select(BenSampID,  RepNum, Year, Season, SCI) %>%
  arrange(Year, Season)
# Site data for mapview
sites_sf <- distinct(IR_SCI, StationID, .keep_all = T) %>%
  filter(!is.na(Lat) | !is.na(Long)) %>% # get rid of missing lat/long
  select(BenSampID, StationID, Lat, Long, StreamName, Basin,
         Location, Region, EcoRegion, EcoregionCode, County) %>%# drop data to avoid any confusion
  st_as_sf(coords = c("Long", "Lat"),  # make spatial layer using these columns
           remove = T, # don't remove these lat/lon cols from df
           crs = 4326)
# IR decision
IR_decision <- select(IR_SCI, BenSampID, StationID, IRYear:FinalAssessmentRating)


shared_IR_SCI <- SharedData$new(IR_SCI, ~BenSampID, group = 'Choose Station' )
shared_IR_HAB <- SharedData$new(IR_HAB, ~BenSampID, group = 'Choose Station' )
shared_IR_HAB_heatmap <- SharedData$new(IR_HAB_heatmap, ~BenSampID, group = 'Choose Station' )
shared_IR_SCI_Table <- SharedData$new(IR_SCI_Table, ~BenSampID, group = 'Choose Station' )
shared_IR_SCI_Raw <- SharedData$new(IR_SCI_Raw, ~BenSampID, group = 'Choose Station' )
shared_sites_sf <- SharedData$new(sites_sf, ~BenSampID, group = 'Choose Station' )
shared_IR_decision <- SharedData$new(IR_decision, ~BenSampID, group = 'Choose Station' )
shared_IR_SCI_HAB2 <- SharedData$new(IR_SCI_HAB2, ~BenSampID, group = 'Choose Station' )


bscols(
  list(
    filter_checkbox('CollectorID', "Collector ID", shared_IR_SCI, ~CollectorID, inline = TRUE),
    filter_select("Basin", "Basin", shared_IR_SCI, ~Basin),
    filter_select("StationID", "Station ID", shared_IR_SCI, ~StationID),
    filter_checkbox('RepNum', "Replicate Numbers to Include in Calculations (if none are selected then all are included)", shared_IR_SCI, ~RepNum, inline = TRUE)

  )
)
   
```

Notice anything weird with the Basin filter? That's because of all the errors in the database. If you can't easily find your data by basin, that means it needs to be fixed. When you are completing the IR2020 form it may be a good idea to fix your errors.




SCI Scores
=====================================  

Column {data-width=200}
-------------------------------------

### SCI Plot

```{r bio info tabset, echo=FALSE}
plot_ly(shared_IR_SCI, x = ~CollDate, y = ~`SCI`, type = 'bar', 
        color = ~Season, width = .10,
        #marker = list(line = list(width = 1.5)),
            hoverinfo="text", text=~paste(sep="<br>",
                                      paste("StationID: ", StationID),
                                      paste("Collection Date: ", CollDate),
                                      paste("Collector ID: ",CollectorID),
                                      paste("SCI Score: ", `Fam SCI`))) %>%
      add_lines(x= ~ CollDate,y= ~ limit, mode='line',line = list(color = 'red'),
                hoverinfo = "text", text="SCI Limit", name="SCI Limit") %>%
      layout(#showlegend=FALSE,
             yaxis=list(title="SCI"),
             xaxis=list(title="Sample Date",tickfont = list(size = 10)))

```

### SCI Table

```{r table, echo=FALSE}

  #div( height="400px", background_color="gray")),
  datatable(shared_IR_SCI_Table,extensions = 'Buttons',  rownames = FALSE,
            options=list(dom= 'Bt', deferRender=TRUE, scrollY=250,scroller=TRUE, buttons=list('copy'))) # this table does update and filter on 'method' selection

```



Habitat Scores
=====================================  

Column {data-width=200}
-------------------------------------

### Total Habitat Plot

```{r tothab plot, echo=FALSE}

box1 <- data.frame(CollDate = c(min(IR_HAB$CollDate) - months(1), 
                                  min(IR_HAB$CollDate) - months(1), 
                                  max(IR_HAB$CollDate) + months(1),
                                  max(IR_HAB$CollDate) + months(1)), 
                   y = c(0, 100, 100, 0))
box2 <- data.frame(CollDate = c(min(IR_HAB$CollDate) - months(1), 
                                  min(IR_HAB$CollDate) - months(1), 
                                  max(IR_HAB$CollDate) + months(1),
                                  max(IR_HAB$CollDate) + months(1)), 
                   y = c(100, 130, 130, 100))
box3 <- data.frame(CollDate = c(min(IR_HAB$CollDate) - months(1), 
                                  min(IR_HAB$CollDate) - months(1), 
                                  max(IR_HAB$CollDate) + months(1),
                                  max(IR_HAB$CollDate) + months(1)), 
                   y = c(130, 150, 150, 130))
box4 <- data.frame(CollDate = c(min(IR_HAB$CollDate) - months(1), 
                                  min(IR_HAB$CollDate) - months(1), 
                                  max(IR_HAB$CollDate) + months(1),
                                  max(IR_HAB$CollDate) + months(1)), 
                   y = c(150, 200, 200, 150)) 

plot_ly(shared_IR_HAB) %>%
  add_polygons(x = ~CollDate, y = ~y, data = box1, fillcolor = "firebrick",opacity=0.6, line = list(width = 0),
                   hoverinfo="text", name =paste('High Probability of Stress to Aquatic Life')) %>%
  add_polygons(data = box2, x = ~CollDate, y = ~y, fillcolor = "#F0E442",opacity=0.6, line = list(width = 0),
                   hoverinfo="text", name =paste('Medium Probability of Stress to Aquatic Life')) %>%
  add_polygons(data = box3, x = ~CollDate, y = ~y, fillcolor = "#009E73",opacity=0.6, line = list(width = 0),
                   hoverinfo="text", name =paste('Low Probability of Stress to Aquatic Life')) %>%
  add_polygons(data = box4, x = ~CollDate, y = ~y, fillcolor = "#0072B2",opacity=0.6, line = list(width = 0),
                   hoverinfo="text", name =paste('No Probability of Stress to Aquatic Life')) %>%
  add_lines(data = shared_IR_HAB, x = ~CollDate, y = ~hab1, line = list(color = "firebrick"),
            hoverinfo="text", name ='High Probability of Stress to Aquatic Life',
            text='High Probability of Stress to Aquatic Life') %>%
  add_lines(data = shared_IR_HAB, x = ~CollDate, y = ~hab2, line = list(color = "#F0E442"),
            hoverinfo="text", name ='Medium Probability of Stress to Aquatic Life',
            text = 'Medium Probability of Stress to Aquatic Life') %>%
  add_lines(data = shared_IR_HAB, x = ~CollDate, y = ~hab3, line = list(color = "#009E73"),
            hoverinfo="text", name ='Low Probability of Stress to Aquatic Life',
            text = 'Low Probability of Stress to Aquatic Life' ) %>%
  add_lines(data = shared_IR_HAB, x = ~CollDate, y = ~hab4, line = list(color = "#0072B2"),
            hoverinfo="text", name = 'No Probability of Stress to Aquatic Life',
            text = 'No Probability of Stress to Aquatic Life') %>%
  add_bars(data = shared_IR_HAB, x = ~CollDate, y = ~TotalHabitat, type = 'bar', color = ~Season,
            hoverinfo="text", text=~paste(sep="<br>", width = .10,
                                      paste("StationID: ", StationID),
                                      paste("Collection Date: ", CollDate),
                                      paste("Collector ID: ", CollectorID),
                                      paste("Total Habitat Score: ", TotalHabitat))) %>%
  layout(showlegend=FALSE,
             yaxis=list(title="SCI"),
             xaxis=list(title="Sample Date",tickfont = list(size = 10)))

```



### RBP Habitat Heatmap

```{r total hab heatmap, echo = FALSE}
# Set how many colors you will use and call them out by hex name
brks <- 1:19
clrs <- c("#8B0000", "#9D0000", "#AF0000", "#C10000", "#D40000", "#E60000", "#F80000", "#FF1415", "#FF3235", "#FF5055", "#FF6F75",
          "#FF8D95", "#FFABB5", "#FFC3CD", "#FFCDD5", "#FFD7DE", "#FFE1E6", "#FFEBEE", "#FFF5F6", "#FFFFFF")

datatable(shared_IR_HAB_heatmap, escape = F, rownames = F, extensions = 'Buttons',
          colnames = c('BenSampID','HabSampID','Collection Date','Enter Date',
                       'Comments','Channel Alteration','Banks', 
                       'Bank Vegetation', 'Cover', 'Embeddedness', 'Flow', 
                       'Pool Substrate', 'Pool Variability', 'Riffles',
                       'Riparian Vegetation', 'Sediment',  'Sinuosity',
                       'Substrate','Velocity', 'Total Habitat'),
          options = list(dom='Bt', buttons=list('copy'))) %>%
  formatStyle(c('ALTER', 'BANKS', 'BANKVEG', 'COVER', 'EMBED', 'FLOW', 'POOLSUB', 'POOLVAR', 
      'RIFFLES', 'RIPVEG', 'SEDIMENT', 'SINUOSITY', 'SUBSTRATE', 'VELOCITY', 'TotalHabitat'),
    backgroundColor = styleInterval(brks, clrs), 
              textAlign = 'center', `font-family` = 'Arial') %>%
  formatStyle(c('ALTER', 'BANKS', 'BANKVEG', 'COVER', 'EMBED', 'FLOW', 'POOLSUB', 'POOLVAR', 
      'RIFFLES', 'RIPVEG', 'SEDIMENT', 'SINUOSITY', 'SUBSTRATE', 'VELOCITY', 'TotalHabitat'), 
      fontWeight = styleInterval(10, c('bold','normal')), 
              textAlign = 'center', `font-family` = 'Arial') %>%
  formatStyle('TotalHabitat', backgroundColor = "lightgray")

```



Station Summary {data-orientation=rows}
=====================================  

Row {data-height=325}
-------------------------------------


### Bio Default Stats

```{r default stats, echo=FALSE}

span(strong('Note: Averages inlcude ALL replicate samples.'), 'If you wish to not include replicates in an average, you will need to use the Bio Interactive Stats feature or calculate by hand.')

datatable(shared_IR_SCI_HAB2, extensions = 'Buttons',  rownames = FALSE, 
          options = list(dom= 'Bt', buttons=list('copy'),
                         pageLength = 1,
                         columnDefs = list(list(visible=FALSE, targets=c(0,1))))) # hide BenSampID column, note DT is JavaScript library and JS starts counting at 0 so colnumber needs to be 0 to work
```



Row
-------------------------------------


### Bio Interactive Stats

```{r yearly Avg, echo=FALSE}
filter_checkbox("year", "Choose Year(s) to Average (if none are selected all are used by default)", shared_IR_SCI, ~Year, inline = TRUE)
span('Selected Year(s) Average SCI : ',
     summarywidget(shared_IR_SCI, statistic='mean', column='SCI', #selection=~Year, 
              digits=1))
```

<br>

Selected Year(s) Fall Average SCI :
```{r IR Fall Avg, echo=FALSE}
summarywidget(shared_IR_SCI, statistic='mean', column='SCI', selection=~Season=='Fall', digits=1)
```

<br>

Selected Year(s) Spring Average SCI :
```{r IR Spring Avg, echo=FALSE}
summarywidget(shared_IR_SCI, statistic='mean', column='SCI', selection=~Season=='Spring', digits=1)
```


### Bio Interactive Results

```{r SCI Raw, echo=FALSE}
datatable(shared_IR_SCI_Raw, extensions = 'Buttons',  rownames = FALSE, options = list(dom= 'Bt', buttons=list('copy')))
```


Row
-------------------------------------


### Site Map

```{r site map, echo=FALSE}

#mapview(shared_sites_sf, label= sites_sf$StationID) # mapview doesnt work with crosstalk, yet

CreateWebMap(maps = c("Topo","Imagery","Hydrography"), collapsed = TRUE) %>%
  setView(-78, 37.5, zoom=6) %>%
  addCircleMarkers(data=shared_sites_sf,color='yellow', fillColor='blue', radius = 5,
                   fillOpacity = 0.5,opacity=1,weight = 2,stroke=T,group="sites",
                   label = ~StationID,
                   popup=~paste(sep='<br>',
                                paste(strong('StationID : '), StationID),
                                paste(strong('Location :'), Location),
                                paste(strong('Stream Name : '), StreamName),
                                paste(strong('Basin : '), Basin),
                                paste(strong('DEQ Region : '), Region),
                                paste(strong('County : '), County),
                                paste(strong('Ecoregion Code : '), EcoregionCode),
                                paste(strong('EcoRegion : '), EcoRegion))) %>%
  inlmisc::AddHomeButton(raster::extent(-83.89, -74.80, 36.54, 39.98), position = "topleft") %>%
  inlmisc::AddSearchButton(group = "sites", zoom = 15,propertyName = "label",
                           textPlaceholder = "Search stations") %>%
  addLayersControl(baseGroups=c("Topo","Imagery","Hydrography"),
                       overlayGroups = c('sites'),
                       options=layersControlOptions(collapsed=T),
                       position='topleft')
  


```



Assessment Decision {data-orientation=columns}
=====================================  

Column {data-width=200}
-------------------------------------

### Assessment Decision

```{r assessmentDecision, echo=FALSE}
datatable(shared_IR_decision, extensions = 'Buttons',  rownames = FALSE, width="100%", 
          options = list( columnDefs = list(list(className = 'dt-center', targets = 4)),
                           pageLength = 15,
                           lengthMenu = c(5, 10, 15, 20),
                          dom= 'Bftlir', 
                          buttons = list('copy', 'print', 
                                         list(extend = 'collection',
                                              buttons = c('csv', 'excel', 'pdf'),
                                              text = 'Download' ))))
```

